---
title: "analysis_AayushGaywala"
output: html_document
date: '2022-07-24'
---

Aayush Gaywala
Mr. Jose Aveldanes
Section #1
Data Science w/ R and Python
Final Project - Linear Regression Analysis

# Description of the Model

Data Question: What is the relationship between fertility and examination (fitness)?

Modeling Objective: To predict fertility given examination.

Data: This data set, swiss, is a built-in data set of R that has 47 observations on 6 variables. When Switzerland was developing and in a demographic transition (1888), this survey took data, such as fertility measure and other socio-economic factors, from 47 French-speaking "provinces". 

Response Variable: Standardized Fertility Measure

The model used in this analysis is a linear regression model. It is a very simple form of supervised machine learning.
I chose linear regression for this specific model because this relationship can be modeled well with a straight line.
I have not used interaction terms in my model.

# Exploratory Data Analysis

Exploratory Data Analysis is just learning about your data and the relationships within it.

To run most of the code below, one needs to install the following packages: dplyr, ggplot2, and ggpubr.

```{r}
library(dplyr)
library(ggplot2)
library(ggpubr)
```

First, we will just get to know the data set and its specifications.

```{r}
head(swiss)
names(swiss)
dim(swiss)
```

Then, we can look at the distributions for each of the variables.

```{r}
par(mfrow = c(3, 2))
hist(swiss$Fertility)
hist(swiss$Agriculture)
hist(swiss$Examination)
hist(swiss$Catholic)
hist(swiss$Education)
hist(swiss$Infant.Mortality)
```

Above you can see that our concerned variables (examination and fertility) are generally normally distributed.

By using the correlation function, we learn that there is negative relationship between both of them.

```{r}
cor(swiss$Fertility, swiss$Examination)
```

The below code chunk creates a scatterplot with a line of best fit and the equation in order to help the reader visualize the relationship between the data.

The reader should note that the y-axis does not start from zero. It instead starts from ~ 30.

Another important note is that the below graph utilizes all of the data, not just the train and test data sets.

```{r}
lm.graph <- swiss %>% ggplot(aes(x = Examination, y = Fertility)) + geom_point(size = 5, alpha = 0.75) + geom_smooth(method = "lm", se = F, col = "blue") + stat_regline_equation(label.x = 30, label.y = 88) + ggtitle("Examination vs. Fertility")
print(lm.graph)
```

# Output of the Final Model

By setting the seed, we can guarantee re-producible results.

Then, partition the data (80:20 split) in order to have a train data set (to train our model off of) and a test data set (for predicting after the model has been created). 

```{r}
set.seed(47)
index <- sample(1:nrow(cars), 0.8*nrow(cars))
train_swiss <- swiss[index, ]
test_swiss <- swiss[-index, ]
```

By using base R, I made a linear regression model and then used summary() to summarize the model and find the p-values, coefficients, intercepts, etc.

Running this code chunk will give you those statistics.

```{r}
linear_regression <- lm(Fertility~Examination, data = train_swiss)
summary(linear_regression)
```

Once we have trained the model, we can now use it to predict our test data values.

```{r}
predictions <- predict(linear_regression, test_swiss)
print(predictions)
```

We can now combine the actuals and the predictions into a data frame and calculate multiple statistics such as Min-Max Accuracy and Mean Absolute Percentage Error. 

```{r}
actualsAndPredicteds <- data.frame(cbind(actuals = test_swiss$Fertility, predicteds = predictions))
actualsAndPredicteds
```

Below is a code chunk that calculates some statistics for error.

``` {r}
# Min-Max Accuracy

minMaxAcc <- mean(apply(actualsAndPredicteds,1, min)/apply(actualsAndPredicteds,1, max))

minMaxAcc

# Mean Absolute Percentage Error
mape <- mean(abs((actualsAndPredicteds$predicteds - actualsAndPredicteds$actuals))/actualsAndPredicteds$actuals)
 
mape 
```
Min Max Accuracy being about 0.9 is quite good, and Mean Absolute Percentage Error being about 0.1 is also quite good.

Conclusion: 

By using the model, I have found a weak relationship between Examination and Fertility rates (p-value < 0.001, Adjusted R-squared - 0.3778). This relationship is a negative one with y-intercept at (0, 86.4372) and a decrease of 0.99 in Fertility for every increase in 1 Examination unit. 

# Discussion of the Assumptions for the Final Model

There are 4 general assumptions for every linear regression model.

Independence of Observations:
- This depends on if Examination and Fertility as statistics were collected independently of each other and were collected using valid statistical methods. 
- We can only hope that the R Datasets package collected data properly.

Normality (following a indepedent distribution):
- We can check if our dependent and independent variables roughly follow a normal distribution.

```{r}
par(mfrow = c(1, 2))
hist(swiss$Fertility)
hist(swiss$Examination)
```

- Both are very roughly normal distributed. What matters is the fact that most of the data is concentrated in the middle.

Linearity:
- This is whether or not the relationship between the variables can be represented by a linear equation.

```{r}
plot(Fertility~Examination, data = swiss)
```

- Yes! It does looks like the data can be represented through a linear equation.

Homogeneity of variance:
- This is a test to see if the size of error of our prediction doesn't change a lot over the range of our independent variable.

```{r}
par(mfrow = c(2, 2))
plot(linear_regression)
```

- As we can see from the plot, the residuals stay around the center line most of the time, so we can see that this assumption is true.

# Interpretations of the Coefficents

The correlation between physical ability and fertility rates seems rather odd at first.
However, the relationship can be explained through science. When a female/male does physical activity, hormones related to reproductive systems may be suppressed which causes infertility.
Seeing as this data was collected in 1888 and that 1888 was a much different time, this relationship can be explained by a multitude of things.

# Additional Work

As the final model is a result of editing and refining the models, I didn't save previous in progress models. Therefore, no additional work is present.
